---
title: "P8105_hw2_dry2115"
author: "Dayoung Yu"
date: 2018-10-05
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


## Problem 1

Read in and clean nyc transit data

```{r}
transit_df = 
  read_csv(file = "./data/nyc_transit_data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), 
        entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The data set _transit df_ contains information on entrance specifications as well as available routes and ADA compliancy for each subway station in New York City. The variables included are: `r names(transit_df)`. These variables were specifically selected from the parent NYC_Transit_Subway_Entrance_And_Exit_Data dataset. The janitor function was used to clean the variable names and the entry variable was converted from character type to logical. The dimensions of _transit df_ are `r ncol(transit_df)` columns by `r nrow(transit_df)` rows. These data are not tidy because the route1-route11 columns should be collapsed into one column.



```{r}
stations_df = distinct(transit_df, station_name, line, .keep_all = TRUE)
  distinct_stations = nrow(stations_df)
  ada_compliant = nrow(filter(stations_df, ada == TRUE))
  
  without_vending = nrow(filter(stations_df, vending == "NO"))
  entrance_without_vending = nrow(filter(stations_df, vending == "NO", entry == TRUE))
  prop_entrance = round((entrance_without_vending / without_vending), digits = 3)
  
```

There are `r distinct_stations` distinct stations in the data set. `r ada_compliant` stations are ADA compliant. The proportion of non-vending stations that allow entry is `r prop_entrance`.


```{r}
transit_tidy_df = gather(transit_df, key = route_number, value = route_name, route1:route11)

  a_stations = 
    filter(transit_tidy_df, route_name == "A") %>%
    distinct(station_name, line) %>%
    nrow()
  
  compliant_a_stations = 
    filter(transit_tidy_df, route_name == "A", ada == TRUE) %>%
    distinct(station_name, line) %>%
    nrow()
```

`r a_stations` distinct stations serve the A train. Of those stations `r compliant_a_stations` stations are ADA compliant.

## Problem 2

Read and clean Mr. Trash Wheels data:
```{r}
library(readxl)

trash_df = 
  read_excel("./data/trashwheel_data_updated.xlsx", sheet = 1, range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  filter(dumpster != is.na(dumpster)) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls, digits = 0))
  )

```

Read and clean precipation data:
```{r}
prcp16_df = (
  read_excel("./data/trashwheel_data_updated.xlsx", sheet = 5, range = cell_rows(2:14)) %>%
  janitor::clean_names() %>%
  rename(total_prcp = total) %>%
  mutate(year = 2016))

prcp17_df = (
  read_excel("./data/trashwheel_data_updated.xlsx", sheet = 4, range = cell_rows(2:14)) %>%
  janitor::clean_names() %>%
  rename(total_prcp = total) %>%
  mutate(year = 2017))

```

Combine precipation data from 2016 and 2017:

```{r}

```


## Problem 3

Install p8105 data package:
```{r}
devtools::install_github("p8105/p8105.datasets")
```


```{r}

```

