---
title: "P8105_hw2_dry2115"
author: "Dayoung Yu"
date: 2018-10-05
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


## Problem 1

Read in and clean nyc transit data

```{r}
transit_data = 
  read_csv(file = "./data/nyc_transit_data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), 
        entry, vending, starts_with("entrance"), ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The data set _transit data_ contains information on entrance specifications as well as available routes and ADA compliancy for each subway station in New York City. The variables included are: `r names(transit_data)`. These variables were specifically selected from the parent NYC_Transit_Subway_Entrance_And_Exit_Data dataset. The janitor function was used to clean the variable names and the entry variable was converted from character type to logical. The dimensions of _transit data_ are `r ncol(transit_data)` columns by `r nrow(transit_data)` rows. These data are not tidy because the route1-route11 columns should be collapsed into one column.



## Problem 2

```{r}

```




# Problem 3

install p8105 data package
```{r}
devtools::install_github("p8105/p8105.datasets")
```


```{r}

```

